---
title: "Identification of Heterogeneous Causal Effects"
author: "Falco J. Bargagli-Stoffi"
date: "14/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Volumes/GoogleDrive/My Drive/Research//BCF-IV/Data')
```

```{r include=FALSE}
# Load packages
rm(list=ls())
library(rpart.utils)
library(rpart.plot)
library(rdrobust)
library(haven)
library(bcf)
library(AER)
source("bcf_iv.R")
```

# Identifying the Causal Effects

In this section we estimate the causal effects of additional funding on students performances.
In particular, our outcome variables are \texttt{certificate} (which is a dummy variable that assumes value 1 if the student gets an "A" certificate: the highest possible grade in the Flanders) and \texttt{progress school} (which is a dummy recording whether or not the student got school retention).

The following analyses were performed using the algorithm proposed in the paper "Heterogeneous causal effects with imperfect compliance: a novel Bayesian machine learning approach" by Falco J. Bargagli Stoffi, Kristof De Witte and Giorgio Gnecco.

```{r include = FALSE}
# Load Data
students_data <- read_dta("first_stage_students_2011.dta")
students_data$eligible_dummy <- students_data$D


myvars <- c("llncode", "schooljaar", "school", "GOKpercentage", "eligible_dummy",
            "GOKschool","PA", "progress_school", "certificate", "ST", "TNN",
            "Opl", "thuisloos","trekkend", "change_school",
            "primary_retention", "man", "BULO", "GONschool","leerkracht_fulltime",
            "leerkracht_diploma", "leerkracht_age", "leerkracht_seniority",
            "leerkracht_female", "directie_age", "leerkracht_unexperienced0_10",
            "directie_unexperienced0_10", "directie_seniority")

students_data <- students_data[myvars]
students_data_2011 <- students_data[which(students_data$schooljaar == 2011),]
table(students_data_2011$GOKschool, students_data_2011$eligible_dummy)
```

# First Outcome: Certificate

We now focus on the effects when we sample units in a bandwidth of 0.035 (which is the optimal bandwidth for the outcome \texttt{certificate}) around the threshold (10\%).

```{r }
students_data_randomized_03_2011 <- 
  students_data_2011[which(students_data_2011$GOKpercentage >= .065 
                           & students_data_2011$GOKpercentage <= .135),]
```

Moreover, from every school we sample a number students in order to increase the balance in the covariate and to guarantee an equal representation to all the schools, avoiding biases related to the over-representation of biggest schools' students.

In the first case, we sample 50 students from each school.

```{r}
schools <- 
students_data_randomized_03_2011$school[which(
!duplicated(students_data_randomized_03_2011$school))]

sample_students <- as.data.frame(matrix(data = NA, nrow = 50*length(schools),
                                        ncol = ncol(students_data_randomized_03_2011)))
colnames(sample_students) <- colnames(students_data_randomized_03_2011)

for (j in (0:(length(schools)-1))){
  set.seed(j  + 2020)
  sample_students[(1+(j*50)):(50+(j*50)),]  <-
  students_data_randomized_03_2011[which(
  students_data_randomized_03_2011$school %in%
  schools[j+1]),][sample(1:nrow(students_data_randomized_03_2011[which(
  students_data_randomized_03_2011$school %in% schools[j+1]),]),
  50, replace = FALSE ),]
}

sample_student <- round(sample_students[, -(1:4)], 0)
sample_students <- cbind(sample_students[,1:4], sample_student)
```

Then, we run our BCF-IV algorithm on this sample of units.

```{r results = 'hide', include = TRUE, message = FALSE, warning = FALSE}
# Attaching the Sample and the Covariates
attach(sample_students)
x <- cbind(primary_retention , man , BULO, 
           leerkracht_age , leerkracht_seniority,
           leerkracht_diploma,
           directie_age, directie_seniority
)
z <- as.matrix(eligible_dummy)
y <- as.matrix(certificate)
w <- as.matrix(GOKschool)
detach(sample_students)
```

Below we run the BCF-IV algorithm.

```{r , warning = FALSE}
set.seed(2020)
bcf_iv(y, w, z, x, max_depth = 2, n_burn = 1000, n_sim = 1000,
       inference_ratio = 0.50, binary = TRUE )
```

As a robustness check, we run the BCF-ITT algorithm.

```{r , warning = FALSE}
set.seed(2020)
bcf_itt(y, w, z, x, max_depth = 2, n_burn = 1000, n_sim = 1000,
        inference_ratio = 0.50, binary = TRUE ) 
```

In the second case we sample 62 students from every school (where 62 units is the size of the smallest school).

```{r}
schools <- 
students_data_randomized_03_2011$school[which(
!duplicated(students_data_randomized_03_2011$school))]

sample_students <- as.data.frame(matrix(data = NA, nrow = 62*length(schools),
                                        ncol = ncol(students_data_randomized_03_2011)))
colnames(sample_students) <- colnames(students_data_randomized_03_2011)

for (j in (0:(length(schools)-1))){
  set.seed(j  + 2020)
  sample_students[(1+(j*62)):(62+(j*62)),]  <-
  students_data_randomized_03_2011[which(
  students_data_randomized_03_2011$school %in%
  schools[j+1]),][sample(1:nrow(students_data_randomized_03_2011[which(
  students_data_randomized_03_2011$school %in% schools[j+1]),]),
  62, replace = FALSE ),]
}

sample_student <- round(sample_students[, -(1:4)], 0)
sample_students <- cbind(sample_students[,1:4], sample_student)
```

```{r results = 'hide', include = TRUE, message = FALSE, warning = FALSE}
# Attaching the Sample and the Covariates
attach(sample_students)
x <- cbind(primary_retention , man , BULO, 
           leerkracht_age , leerkracht_seniority,   
           directie_age, directie_seniority
)
z <- as.matrix(eligible_dummy)
w <- as.matrix(GOKschool)
y <- as.matrix(certificate)
detach(sample_students)
```

Below we run the BCF-IV algorithm.

```{r , warning = FALSE}
set.seed(2020) 
bcf_iv(y, w, z, x, max_depth = 2, n_burn = 1000, n_sim = 1000,
       inference_ratio = 0.50, binary = TRUE )
```

As a robustness check, we run the BCF-ITT algorithm.

```{r , warning = FALSE}
set.seed(2020)
bcf_itt(y, w, z, x, max_depth = 2, n_burn = 1000, n_sim = 1000,
        inference_ratio = 0.50, binary = TRUE )
```

# Second Outcome: School Progress

We now focus on the effects when we sample units in a bandwidth of 0.035 (which is the optimal bandwidth for the outcome \texttt{progress\_school}) around the threshold (10\%).

```{r}
students_data_randomized_03_2011 <- 
  students_data_2011[which(students_data_2011$GOKpercentage >= .063 
                           & students_data_2011$GOKpercentage <= .137),]
```

Moreover, from every school we sample a number students in order to increase the balance in the covariate and to guarantee an equal representation to all the schools, avoiding biases related to the over-representation of biggest schools' students.
In the first case, we sample 50 students from each school.

```{r}
schools <- 
students_data_randomized_03_2011$school[which(
!duplicated(students_data_randomized_03_2011$school))]

sample_students <- as.data.frame(matrix(data = NA, nrow = 50*length(schools),
                                        ncol = ncol(students_data_randomized_03_2011)))
colnames(sample_students) <- colnames(students_data_randomized_03_2011)

for (j in (0:(length(schools)-1))){
  set.seed(j  + 2020)
  sample_students[(1+(j*50)):(50+(j*50)),]  <-
  students_data_randomized_03_2011[which(
  students_data_randomized_03_2011$school %in%
  schools[j+1]),][sample(1:nrow(students_data_randomized_03_2011[which(
  students_data_randomized_03_2011$school %in% schools[j+1]),]),
  50, replace = FALSE ),]
}

sample_student <- round(sample_students[, -(1:4)], 0)
sample_students <- cbind(sample_students[,1:4], sample_student)
```

Then, we run our BCF-IV algorithm on this sample of units.

```{r results = 'hide', include = TRUE, message = FALSE, warning = FALSE}
# Attaching the Sample and the Covariates
attach(sample_students)
x <- cbind(primary_retention , man , BULO, 
           leerkracht_age , leerkracht_seniority,   
           directie_age, directie_seniority
)
z <- as.matrix(eligible_dummy)
w <- as.matrix(GOKschool)
y <- as.matrix(progress_school)
detach(sample_students)
```

Below we run the BCF-IV algorithm.

```{r , warning = FALSE}
set.seed(2020) 
bcf_iv(y, w, z, x, max_depth = 2, n_burn = 1000, n_sim = 1000,
       inference_ratio = 0.50, binary = TRUE )
```

As a robustness check, we run the BCF-ITT algorithm.

```{r , warning = FALSE}
set.seed(2020)
bcf_itt(y, w, z, x, max_depth = 2, n_burn = 1000, n_sim = 1000,
        inference_ratio = 0.50, binary = TRUE )
```

In the second case we sample 62 students from every school (where 62 units is the size of the smallest school).

```{r}
schools <- 
students_data_randomized_03_2011$school[which(
!duplicated(students_data_randomized_03_2011$school))]

sample_students <- as.data.frame(matrix(data = NA, nrow = 62*length(schools),
                                        ncol = ncol(students_data_randomized_03_2011)))
colnames(sample_students) <- colnames(students_data_randomized_03_2011)

for (j in (0:(length(schools)-1))){
  set.seed(j  + 2020)
  sample_students[(1+(j*62)):(62+(j*62)),]  <-
  students_data_randomized_03_2011[which(
  students_data_randomized_03_2011$school %in%
  schools[j+1]),][sample(1:nrow(students_data_randomized_03_2011[which(
  students_data_randomized_03_2011$school %in% schools[j+1]),]),
  62, replace = FALSE ),]
}

sample_student <- round(sample_students[, -(1:4)], 0)
sample_students <- cbind(sample_students[,1:4], sample_student)
```

```{r results = 'hide', include = TRUE, message = FALSE, warning = FALSE}
# Attaching the Sample and the Covariates
attach(sample_students)
x <- cbind(primary_retention , man , BULO, 
           leerkracht_age , leerkracht_seniority,   
           directie_age, directie_seniority
)
z <- as.matrix(eligible_dummy)
w <- as.matrix(GOKschool)
y <- as.matrix(progress_school)
detach(sample_students)
```

Below we run the BCF-IV algorithm.

```{r , warning = FALSE}
set.seed(2020)
bcf_iv(y, w, z, x, max_depth = 2, n_burn = 1000, n_sim = 1000,
       inference_ratio = 0.50, binary = TRUE )
```

As a robustness check, we run the BCF-ITT algorithm.

```{r , warning = FALSE}
set.seed(2020)
bcf_itt(y, w, z, x, max_depth = 2, n_burn = 1000, n_sim = 1000,
        inference_ratio = 0.50, binary = TRUE )
```
